Dependent Types for Safe and Secure Web Programming

### 2013-11-18 paper/conclusion.tex
(Hopefully) final revision of paper before IFL2013 submission.

In particular, the framework guarantees that it is not possible for a CGI 
application to
produce an internal server error due to content being written to the remote
host prior to headers. With regard to database access, we may statically
guarantee that library calls are made in the correct order, and calls to
retrieve rows of data are made only when more data is available. Additionally,
by recording any failures in the resource associated with the database effect,
we may ensure that any failures at certain stages in the process may be
correctly dealt with, and not propagate to further library calls---if a
prepared statement fails to compile, for example, the framework guarantees that
it will not be executed, without causing a fatal runtime error or memory access
violation.
%
Enforcing resource usage protocols also guards against common programmer
errors, saving debugging time by identifying errors at compile time.

>>>>

In particular, the framework guarantees that it is not possible for a CGI 
application to
produce an internal server error due to content being written to the remote
host prior to headers. With regard to database access, we may statically
guarantee that library calls are made in the correct order, and calls to
retrieve rows of data are made only when more data is available. Additionally, by encoding desired invariants within operation types, we may gain static guarantees about adherence to resource usage protocols and failure handling.
Enforcing resource usage protocols also guards against common programmer
errors, saving debugging time by identifying errors at compile time.

### 2013-11-18 paper/protocols.tex
(Hopefully) final revision of paper before IFL2013 submission.

User code executes in the \texttt{TaskRunning} state. Several operations, such
as querying the \texttt{POST} and \texttt{GET} variables, are available in this state, alongside
functions to output data to the web page and append data to the response
headers. It is important to note, that at this stage nothing is written to the
page, with the \texttt{output} and \texttt{addHeader} functions instead
modifying the \texttt{CGIInfo} record. This data may then be printed at the end of the
program's execution, in accordance with the resource usage protocol.

@@

\noindent
The program will be run in \texttt{IO}, and starts and finishes with no active
resources.  It reutnrs either list of \texttt{(String, String)} pairs,
representing names and addresses in the database, or an error.
%This example is shown in Figure
%\ref{fig:testsel}.

@@

The program initially attempts to open a connection to the \texttt{people.db}
database. At this point, since the \texttt{OpenDB} operation has been invoked,
the program transitions to the \texttt{ConnectionOpened} state. The
\texttt{openDB} function returns a \texttt{Bool} indicating whether the
operation is successful. If not, the \texttt{connFail} function
generates an appropriate error and disposes of the resources, as shown in
Figure~\ref{fig:sqlitestates}.

>>>>

User code executes in the \texttt{TaskRunning} state. Several operations, such
as querying the \texttt{POST} and \texttt{GET} variables, are available in this state, alongside
functions to output data to the web page and append data to the response
headers. It is important to note that at this stage nothing is written to the
page, with the \texttt{output} and \texttt{addHeader} functions instead
modifying the \texttt{CGIInfo} record. This data may then be printed at the end of the
program's execution, in accordance with the resource usage protocol.

@@

\noindent
The program will be run in \texttt{IO}, and starts and finishes with no active
resources.  It returns either a list of \texttt{(String, String)} pairs,
representing names and addresses in the database, or an error.
%This example is shown in Figure
%\ref{fig:testsel}.

@@

The program initially attempts to open a connection to the \texttt{people.db}
database. At this point, since the \texttt{OpenDB} operation has been invoked,
the program transitions to the \texttt{ConnectionOpened} state. The
\texttt{openDB} function returns either an error code, if the connection fails, or a unit type should the connection succeed, as shown in Figure \ref{fig:sqlitestates}.
